---
title: "Smiley Real Estate Analysis"
author: "Aro Roseman"
date: "2025-12-05"
output: 
  html_document:
    toc: true
    toc_float: true
    section_autonumber: true
    theme: readable
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
```

```{r libraries, include=FALSE}
library(readr)
library(tidyverse)
library(rlang)
library(purrr)
library(plotly)
library(knitr)
library(kableExtra)
library(scales)
```

```{r import, include=FALSE}
agents <- read_csv("C:/Users/alyss/OneDrive/Job applications/Clever Real Estate Working Interview/agents.csv")
calls <- read_csv("C:/Users/alyss/OneDrive/Job applications/Clever Real Estate Working Interview/calls.csv")
customers <- read_csv("C:/Users/alyss/OneDrive/Job applications/Clever Real Estate Working Interview/customers.csv")
markets <- read_csv("C:/Users/alyss/OneDrive/Job applications/Clever Real Estate Working Interview/markets.csv")
sales <- read_csv("C:/Users/alyss/OneDrive/Job applications/Clever Real Estate Working Interview/sales.csv")
```

---

# 1. Data Cleaning and Preparation 🧹
This first section is about cleaning and preparing the data. It's not very
interesting and isn't very nice to look at, so it's hidden from this report.
However, feel free to [view the complete code](Clever Real Estate Working Interview Project.Rmd) to see the full cleaning steps and take a look at the notes I left 
about why I made the corrections I did. But for now, let's get straight to the 
fun part!

```{r data_cleaning, include=FALSE}
str(agents)
str(calls)
str(customers)
str(markets)
str(sales)
```

```{r data_type_correct, include=FALSE}
#Convert datetime character type to datetime object
calls$call_date <- mdy_hm(calls$call_date)
customers$created_at <- ymd_hms(customers$created_at)
sales$sale_date <- mdy_hms(sales$sale_date)

#Remove special characters from market$price_new_list_median and price_closed_median
#and sales$amount and convert to numeric
markets <- markets %>%
  mutate(
    price_new_list_median = str_remove_all(price_new_list_median, "[$,]"),
    price_new_list_median = as.numeric(price_new_list_median)
  )
markets <- markets %>%
  mutate(
    price_closed_median = str_remove_all(price_closed_median, "[$,]"),
    price_closed_median = as.numeric(price_closed_median)
  )
sales <- sales %>%
  mutate(
    amount = str_remove_all(amount, "[$,]"),
    amount = as.numeric(amount)
  )

#Converts ids to factors
agents <- agents %>%
  mutate(agent_id = as.factor(agent_id))
agents <- agents %>%
  mutate(market_id = as.factor(market_id))
calls <- calls %>%
  mutate(call_id = as.factor(call_id))
calls <- calls %>%
  mutate(customer_id = as.factor(customer_id))
customers <- customers %>%
  mutate(customer_id = as.factor(customer_id))
customers <- customers %>%
  mutate(market_id = as.factor(market_id))
markets <- markets %>%
  mutate(market_id = as.factor(market_id))
markets <- markets %>%
  mutate(census_name = as.factor(census_name))
markets <- markets %>%
  mutate(common_name = as.factor(common_name))
sales <- sales %>%
  mutate(sale_id = as.factor(sale_id))
sales <- sales %>%
  mutate(customer_id = as.factor(customer_id))
```

```{r date_logic_check, include=FALSE}
#Define the comparison time
current_time <- as.POSIXct("2025-12-05 12:00:00", tz="UTC") 

# Define the historical cut-off: Anything before January 1, 2025 is an error
historical_cutoff <- as.POSIXct("2025-01-01 00:00:00", tz="UTC")

#calls
calls_future_dates <- calls %>%
  filter(call_date > current_time)

calls_old_dates <- calls %>%
  filter(call_date < historical_cutoff)

#customers
customers_future_dates <- customers %>%
  filter(created_at > current_time)

customers_old_dates <- customers %>%
  filter(created_at < historical_cutoff)

#sales
sales_future_dates <- sales %>%
  filter(sale_date > current_time)

sales_old_dates <- sales %>%
  filter(sale_date < historical_cutoff)

# Print the results
print(calls_future_dates)
print(calls_old_dates)

print(customers_future_dates)
print(customers_old_dates)

print(sales_future_dates)
print(sales_old_dates)
```

```{r missing_value_check, include=FALSE}
#Number of missing values
agents_na <- agents %>%
  summarise(across(everything(), ~sum(is.na(.))))
calls_na <- calls %>%
  summarise(across(everything(), ~sum(is.na(.))))
customers_na <- customers %>%
  summarise(across(everything(), ~sum(is.na(.))))
markets_na <- markets %>%
  summarise(across(everything(), ~sum(is.na(.))))
sales_na <- sales %>%
  summarise(across(everything(), ~sum(is.na(.))))

#Percentage of missing values
na_percent <- function(df) {
  sapply(df, function(x) round(sum(is.na(x)) / length(x) * 100, 2))
}
na_percent_agents <- na_percent(agents)
print(na_percent_agents)
na_percent_calls <- na_percent(calls)
print(na_percent_calls)
na_percent_customers <- na_percent(customers)
print(na_percent_customers)
na_percent_markets <- na_percent(markets)
print(na_percent_markets)
na_percent_sales <- na_percent(sales)
print(na_percent_sales)

#Check missing values
rows_with_na_agents <- agents %>%
  filter(rowSums(is.na(.)) > 0)
print(rows_with_na_agents)
#Low missingness - impute missing values with median
median_customer_reviews_2024 <- median(agents$customer_reviews_2024, na.rm = TRUE)
agents <- agents %>%
  replace_na(list(customer_reviews_2024 = median_customer_reviews_2024))
median_avg_rating_2024 <- median(agents$avg_rating_2024, na.rm = TRUE)
agents <- agents %>%
  replace_na(list(avg_rating_2024 = median_avg_rating_2024))

rows_with_na_calls <- calls %>%
  filter(rowSums(is.na(.)) > 0)
print(rows_with_na_calls)
#High percentage of missing values, but only for call handlers of disconnected
#calls or voicemails. These values should be NA and so are not truly missing

rows_with_na_sales <- sales %>%
  filter(rowSums(is.na(.)) > 0)
#sales rows with missing values are missing all values - drop rows
sales <- sales %>%
    filter(!if_all(everything(), is.na))
```

```{r outliers_check, include=FALSE}
check_outliers <- function(df, col_name, k = 1.5) {
  col_sym <- sym(col_name) 

  #Calculate IQR components
  Q1 <- quantile(df[[col_name]], 0.25, na.rm = TRUE)
  Q3 <- quantile(df[[col_name]], 0.75, na.rm = TRUE)
  IQR_val <- Q3 - Q1
  
  lower <- Q1 - (k * IQR_val)
  upper <- Q3 + (k * IQR_val)
  
  #Identify and count the outliers
  outliers <- df %>%
    filter(!!col_sym < lower | !!col_sym > upper) %>%
    # Select only the key columns for review
    select(!!col_sym, everything())
    
  #Return the results as a list
  return(list(
    column = col_name,
    count = nrow(outliers),
    percentage = round(nrow(outliers) / sum(!is.na(df[[col_name]])) * 100, 2),
    outlier_rows = outliers
  ))
}

df_list <- list(
  agents = agents, 
  calls = calls, 
  customers = customers, 
  markets = markets, 
  sales = sales 
)

columns_to_check <- list(
  agents = c("buyer_customers_sent_2024", "seller_customers_sent_2024",
             "buyer_closes_2024", "seller_closes_2024", "customer_reviews_2024",
             "avg_rating_2024"),
  markets = c("population", "days_on_market_median", "price_new_list_median",
              "price_closed_median", "listings_on_market_count", "listings_price_increase_count", "listings_price_decrease_count", "listings_closed_count"),
  sales = c("amount")
)

results <- imap(df_list, function(df, df_name) {
  
  cols <- columns_to_check[[df_name]]
  
  if (is.null(cols)) {
    return(NULL)
  }
  
  outlier_results <- map(cols, ~ check_outliers(df, .x))
  
  return(outlier_results)
})

results <- compact(results)

# Create a summary table of all outlier counts
outlier_summary <- imap_dfr(results, function(df_res, df_name) {
  map_dfr(df_res, ~ data.frame(
    Dataset = df_name,
    Column = .x$column,
    Outlier_Count = .x$count,
    Percent_Outliers = .x$percentage
  ))
})

print(outlier_summary)

#Look at rows with outliers
get_outlier_rows <- function(dataset_name, column_name) {
  result_list <- results[[dataset_name]]
  match_index <- map_lgl(result_list, ~ .x$column == column_name)
  return(result_list[match_index][[1]]$outlier_rows)
}

#customer_reviews_2024
customer_reviews_2024_outliers <- get_outlier_rows("agents", "customer_reviews_2024")
print(customer_reviews_2024_outliers)
#15 is a high but not unreasonable number of reviews. Leave data as is, but use
#caution when interpreting results

#population
population_outliers <- get_outlier_rows("markets", "population")
print(population_outliers)
#All population outliers are large metro areas (New York City, Los Angeles, and
#Chicago). Leave data as is

#price_new_list_median
price_new_list_median_outliers <- get_outlier_rows("markets", "price_new_list_median")
print(price_new_list_median_outliers)
#Outlier new_list_median prices are all in the most expensive cities (Los Angeles,
#San Francisco, and San Jose) according to "Price-to-Rent Ratio: The Best U.S. Cities for Home Buyers and Renters (2024 Data)" by Jaime Dunaway-Seale on Clever Real Estate.
#Prices are credible, so leave data as is, but interpret with caution

#price_closed_median
price_closed_median_outliers <- get_outlier_rows("markets", "price_closed_median")
print(price_closed_median_outliers)
#Outlier closed_median prices are the same expensive cities as price_new_list_median
#outliers. Data is credible, so leave as is

#listings_on_market_count
listings_on_market_count_outliers <- get_outlier_rows("markets", "listings_on_market_count")
print(listings_on_market_count_outliers)
#All outliers are in major metro areas. However, 10% of data are outliers, so
#interpret with extreme caution. Data should be normalized

#listings_price_increase_count
listings_price_increase_count_outliers <- get_outlier_rows("markets", "listings_price_increase_count")
print(listings_price_increase_count_outliers)
#Both outliers are in large metro areas (Houston and Portland) with a high number
#total listings. Outliers appear to be accurate, extreme cases. Data should be
#normalized before interpretation

#listings_price_decrease_count
listings_price_decrease_count_outliers <- get_outlier_rows("markets", "listings_price_decrease_count")
print(listings_price_decrease_count_outliers)
#Outliers are all in large metro areas and likely directly related to large population
#sizes. Leave data as is

#listings_closed_count
listings_closed_count_outliers <- get_outlier_rows("markets", "listings_closed_count")
print(listings_closed_count_outliers)
#Outliers are in large metro areas and likely a direct result of large population.
#Leave data as is

#amount
amount_outliers <- get_outlier_rows("sales", "amount")
print(amount_outliers)
#This is likely a data entry error (this sale is over 52 times higher than the #next largest sale). The customer lives in Detroit, MI where the median price #at close is $205,878. It is likely a decimal point was missed during data #entry, and the correct sale price is $395,928.00. The row will be updated #with this repair
sales <- sales %>%
  mutate(
    amount = ifelse(sale_id == 3458, 395928.00, amount)
  )
print(sales %>% filter(sale_id == 3458))
```

```{r consistency_check, include=FALSE}
#Remove leading and trailing white space and convert all character columns
#to lower case
clean_text <- function(df) {
  df %>%
    mutate(
      across(where(is.character), ~ {
        .x %>%
          trimws() %>%
          tolower()
      })
    )
}

df_list_cleaned <- map(df_list, clean_text)

agents <- df_list_cleaned$agents
calls <- df_list_cleaned$calls
customers <- df_list_cleaned$customers
markets <- df_list_cleaned$markets
sales <- df_list_cleaned$sales
```

```{r normalization, include=FALSE}
#Normalize data affected by population
markets <- markets %>%
  mutate(
    #Percentage of listings that saw a price increase
    increase_rate = listings_price_increase_count / listings_on_market_count,
    
    #Percentage of listings that saw a price decrease
    decrease_rate = listings_price_decrease_count / listings_on_market_count,
    
    #Inventory Turnover Rate
    turnover_rate = listings_closed_count / listings_on_market_count,
    
    #Ratio of Price Change Events to Total Listings (for consistency check)
    price_change_ratio = (listings_price_increase_count + listings_price_decrease_count) / listings_on_market_count
  )
```

```{r normalization_outliers, include=FALSE}
columns_to_check_recheck <- list(
  agents = c("buyer_customers_sent_2024", "seller_customers_sent_2024",
             "buyer_closes_2024", "seller_closes_2024", "customer_reviews_2024",
             "avg_rating_2024"),
  
  markets = c("population", "days_on_market_median", "price_new_list_median",
              "price_closed_median", "listings_on_market_count", "increase_rate",
              "decrease_rate", "turnover_rate"), 
  sales = c("amount")
)

#Update the df_list with the normalized markets df
df_list$markets <- markets

results_normalized <- imap(df_list, function(df, df_name) {
  
  # Use the revised list of columns
  cols <- columns_to_check_recheck[[df_name]]
  
  if (is.null(cols)) {
    return(NULL)
  }
  
  outlier_results <- map(cols, ~ check_outliers(df, .x))
  
  return(outlier_results)
})

results_normalized <- compact(results_normalized)

# Create a summary table of all outlier counts
outlier_summary_normalized <- imap_dfr(results_normalized, function(df_res, df_name) {
  map_dfr(df_res, ~ data.frame(
    Dataset = df_name,
    Column = .x$column,
    Outlier_Count = .x$count,
    Percent_Outliers = .x$percentage
  ))
})

print(outlier_summary_normalized)

#increase_rate outliers
markets_normalized_results <- results_normalized$markets 

increase_rate_outliers <- map_dfr(markets_normalized_results, function(x) {
  if (x$column == "increase_rate") {
    return(x$outlier_rows)
  }
})
print(increase_rate_outliers)
#Portland, Oregon appears to be an extreme case with anomalously high listing
#price increases

#decrease_rate outliers
markets_normalized_results <- results_normalized$markets 

decrease_rate_outliers <- map_dfr(markets_normalized_results, function(x) {
  if (x$column == "decrease_rate") {
    return(x$outlier_rows)
  }
})
print(decrease_rate_outliers)
#Data appears to be accurate but anomalous. New York City is a large, expensive
#market, so price adjustments may not be unusual. Columbus and Indianapolis
#have similar populations and similar median days on market, so they may have 
#faced a systemic market condition (e.g. a market crash in the Midwest) 
```
---

# 2. Analysis and Results 📊
## 2.1 Best Markets 💹
### 2.1.1 Overall Best Markets 🥇
To find the best markets, we'll examine a few different factors. First, let's
see where agents are converting leads into the most sales.

```{r closing_rates, include=FALSE}
agents <- agents %>%
  mutate(
    #Buyer Close Rate
    buyer_close_rate = case_when(
      buyer_customers_sent_2024 == 0 ~ 0, # If no customers sent, rate is 0
      TRUE ~ buyer_closes_2024 / buyer_customers_sent_2024
    ),
    
    #Seller Close Rate
    seller_close_rate = case_when(
      seller_customers_sent_2024 == 0 ~ 0, # If no customers sent, rate is 0
      TRUE ~ seller_closes_2024 / seller_customers_sent_2024
    )
  )
```

```{r market_average_total, include=FALSE}
#Calculate the market averages and totals
agent_market_summary <- agents %>%
  group_by(market_id) %>%
  summarise(
    avg_buyer_close_rate = mean(buyer_close_rate, na.rm = TRUE),
    avg_seller_close_rate = mean(seller_close_rate, na.rm = TRUE),
    total_raw_closes = sum(buyer_closes_2024, na.rm = TRUE) + 
      sum(seller_closes_2024, na.rm = TRUE),
    .groups = 'drop'
  )
```

```{r closing_rate_average_chart}
#Pivot to long format and sort
agent_market_summary_long <- agent_market_summary %>%
  # Calculate an overall average for sorting reference
  mutate(
    overall_avg_rate = (avg_buyer_close_rate + avg_seller_close_rate) / 2
  ) %>%
  #Keep only top 5 markets based on overall average rate
  arrange(desc(overall_avg_rate)) %>%
  slice_head(n = 5) %>%
    #Pivot to long format for ggplot
  pivot_longer(
    cols = starts_with("avg_"),
    names_to = "Close_Rate_Type",
    values_to = "Rate"
  ) %>%
  #Sort markets by overall average rate
  mutate(
    market_id = fct_reorder(market_id, Rate, .fun = mean, .desc = FALSE),
    #Clean up labels for legend/axis
    Close_Rate_Type = case_match(
        Close_Rate_Type,
        "avg_buyer_close_rate" ~ "Buyer Close Rate",
        "avg_seller_close_rate" ~ "Seller Close Rate"
    )
  )

#Create the grouped horizontal bar chart
rate_bar <- ggplot(agent_market_summary_long, aes(x = Rate, y = market_id, fill = Close_Rate_Type,
                                   text = paste("<b>Market:</b>", market_id, "<br>", 
                                                "<b>", Close_Rate_Type, ":</b>", scales::percent(Rate, accuracy = 0.1)))) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(
    aes(x = 0.015, label = paste0("<b>", scales::percent(Rate, accuracy = .1), "</b>")), 
    position = position_dodge(width = 0.9),
    hjust = 0,                                                  
    color = "white",                                            
    size = 3.5
  ) +
  
  #Format x-axis as percentages
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_fill_manual(values = c(
    "Buyer Close Rate" = "#44b2ef", 
    "Seller Close Rate" = "#09223d"
  )) +
  labs(
    title = "<b>Average Buyer and Seller Close Rates by Market<b>",
    x = "Average Close Rate",
    y = "",
    fill = "Transaction Type"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.title.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold")
  )

# Convert to Plotly for interactivity
ggplotly(rate_bar, tooltip = "text") %>%
plotly::layout(legend = list(
    orientation = "h", 
    x = 0.5,             
    y = -0.2,
    xanchor = "center"
  )
  )
```
Based on closing rates alone, it looks like the best markets are Dallas, 
Louisville, Raleigh, Nashville, and Cincinnati. But there's more to the story
than just closing rates. Lots of factors can affect a market's performance - 
let's keep digging!

Other important factors to consider are how long property stays on the market,
the average closing price, the turnover rate, and how often a listing price
increases.

```{r customer_count, include=FALSE}
#Calculate total number of customers per market
customer_market_summary <- customers %>%
  filter(!is.na(market_id)) %>%
  group_by(market_id) %>%
  summarise(
    total_customers = n_distinct(customer_id),
    .groups = 'drop'
  )
```

```{r agents_customers_markets_join, include=FALSE}
#Join summary tables
market_ranking_data <- markets %>%
  select(
    market_id, population, days_on_market_median, price_closed_median, 
    turnover_rate, increase_rate, decrease_rate # Keep decrease_rate for market insights
  ) %>%
  left_join(agent_market_summary, by = "market_id") %>%
  left_join(customer_market_summary, by = "market_id") %>%
  
  # Calculate total customers per 1 million people
  mutate(
    customer_saturation_rate = total_customers / (population / 1000000)
  )
```

When we take all those metrics into account, the best markets seem to be New
York, Los Angeles, Chicago, Dallas, Houston, and Washington, D.C.

```{r unweighted_table}
#Prepare the data
top_markets_table_data <- market_ranking_data %>%
  # Filter to the 5 most populous markets for a meaningful snapshot
  arrange(desc(population)) %>%
  head(5) %>%
  
  # Select the most important metrics
  select(
    market_id,
    population,
    total_customers,
    days_on_market_median,
    price_closed_median,
    increase_rate,
    avg_buyer_close_rate,
    avg_seller_close_rate
  ) %>%
  
  # Rename columns for clarity in the final table
  rename(
    Market = market_id,
    Population = population,
    Customers = total_customers,
    `Median DoM` = days_on_market_median,
    `Median Price` = price_closed_median,
    `Price Increase Rate` = increase_rate,
    `Avg Buyer Close Rate` = avg_buyer_close_rate,
    `Avg Seller Close Rate` = avg_seller_close_rate
  ) %>%
  
  # Format the numbers
  mutate(
    Population = comma(Population, accuracy = 1),
    Customers = comma(Customers, accuracy = 1),
    `Median Price` = dollar(`Median Price`, accuracy = 1),
    `Price Increase Rate` = percent(`Price Increase Rate`, accuracy = 0.1),
    `Avg Buyer Close Rate` = percent(`Avg Buyer Close Rate`, accuracy = 0.1),
    `Avg Seller Close Rate` = percent(`Avg Seller Close Rate`, accuracy = 0.1)
  )

#Generate table
top_markets_table_data %>%
  kbl(caption = "Snapshot of Top 5 Markets by Key Metrics") %>%
  kable_classic_2(full_width = F, html_font = "Cambria") %>%
  row_spec(0, bold = T, background = "#f2f2f2")
```

But, wait! Those are all also large population centers. Let's check where the
best markets are after we control for population. We can standardize each metric
and assign an overall performance score to evaluate it independent of population.

```{r scoring, include=FALSE}
#Select metrics that indicate a "best market"
ranking_metrics <- market_ranking_data %>%
  select(
 #Agent Performance (Higher is better)
 avg_buyer_close_rate, avg_seller_close_rate, total_raw_closes,
 #Market Health (Higher is better)
 turnover_rate, customer_saturation_rate,
 #Market Opportunity (Higher is better)
 price_closed_median, increase_rate,
 #Market Difficulty (Lower is better)
 days_on_market_median
 )

#Standardize all metrics and create Z-scores
standardized_ranking_metrics <- ranking_metrics %>%
 mutate(across(everything(), scale, .names = "Z_{.col}")) %>%
 mutate(Z_days_on_market_median = Z_days_on_market_median * -1) %>% #Invert the Z-score
  
  #Remove the original raw columns before binding
  select(starts_with("Z_")) 

#Calculate composite performance score
market_ranking_final <- market_ranking_data %>%
 bind_cols(standardized_ranking_metrics) %>% 

 mutate(
 #Select all columns that start with "Z_"
 performance_score = rowSums(select(., starts_with("Z_")), na.rm = TRUE)
 ) %>%

 #Rank and select top 5 markets
 arrange(desc(performance_score)) %>%
 mutate(rank = row_number())

# The Top 5 Best Markets
top_5_markets <- market_ranking_final %>%
 head(5)

```

```{r table_weighted}
top_5_score_table_data <- top_5_markets %>%
  
  #Ensure data is sorted by rank derived from the performance_score
  arrange(rank) %>%
  
  # 1. Select the new calculated metrics first, followed by the raw metrics
  select(
    Rank = rank,                             
    Score = performance_score,               
    Market = market_id,
    Population = population,
    Customers = total_customers,
    `Median DoM` = days_on_market_median,
    `Median Price` = price_closed_median,
    `Price Increase Rate` = increase_rate,         
    `Avg Buyer Close Rate` = avg_buyer_close_rate,
    `Avg Seller Close Rate` = avg_seller_close_rate
  ) %>%
  
  # 2. Format the numbers
  mutate(
    Score = round(Score, 2), # Format the calculated score
    Population = comma(Population, accuracy = 1),
    Customers = comma(Customers, accuracy = 1),
    `Median Price` = dollar(`Median Price`, accuracy = 1),
    `Price Increase Rate` = percent(`Price Increase Rate`, accuracy = 0.1),
    `Avg Buyer Close Rate` = percent(`Avg Buyer Close Rate`, accuracy = 0.1),
    `Avg Seller Close Rate` = percent(`Avg Seller Close Rate`, accuracy = 0.1)
  )

#Generate table
top_5_score_table_data %>%
  kbl(caption = "Top 5 Markets by Standardized Composite Performance Score") %>%
  kable_classic_2(full_width = F, html_font = "Cambria") %>%
  row_spec(0, bold = T, background = "#f2f2f2") %>%
  #Highlight Rank and Score columns
  column_spec(1, bold = T, background = "#09223d", color = "white") %>%
  column_spec(2, bold = T, color = "#44b2ef")
```
### 2.1.2 Most Efficient Markets 🚀

California and (and New York, to a lesser extent) still may be overrepresented 
here. The previous scoring was still weighted heavily to the wealthiest and most 
densely populated states, despite normalization. But where is Smiley Real Estate 
most *efficient* rather than merely depending on population and wealth factors? 
The following analysis weights each factor according to importance: closing 
rates and customer saturation are weighted more highly, while median closing 
price is weighted lower to favor efficiency over property value or population.

```{r efficiency}
#Create the Z-score columns
market_ranking_data_scored <- market_ranking_data %>%

  mutate(across(c(avg_buyer_close_rate, avg_seller_close_rate, total_raw_closes, 
                  turnover_rate, customer_saturation_rate, price_closed_median, 
                  increase_rate, days_on_market_median), 
                scale, .names = "Z_{.col}")) %>%
  
  #Invert the Z-score for Days on Market
  mutate(Z_days_on_market_median = Z_days_on_market_median * -1)

#Filter the Z-scores to exclude the raw closes
market_ranking_final_weighted <- market_ranking_data_scored %>%
  
  #Apply new weights to the Z-scores to favor efficiency over wealth and population
  mutate(
    performance_score_weighted = 
      (Z_avg_buyer_close_rate * 1.5) + 
      (Z_avg_seller_close_rate * 1.5) + 
      (Z_customer_saturation_rate * 1.5) + 
      (Z_turnover_rate * 1.0) + 
      (Z_days_on_market_median * 1.0) +
      (Z_price_closed_median * 0.5) + 
      (Z_increase_rate * 1.0)
  ) %>%
  arrange(desc(performance_score_weighted)) %>%
  mutate(rank_weighted = row_number())

# The Top 5 Best Markets For Efficiency
top_5_markets_weighted <- market_ranking_final_weighted %>%
  head(5)
```

```{r table_efficiency}
efficiency_table_data <- top_5_markets_weighted %>%
  
  #Ensure data is sorted by new weighted rank
  arrange(rank_weighted) %>%
  
  #Select the new calculated metrics first, followed by the raw metrics
  select(
    Rank = rank_weighted,                     
    Score = performance_score_weighted,       
    Market = market_id,
    Population = population,
    Customers = total_customers,
    `Median DoM` = days_on_market_median,
    `Median Price` = price_closed_median,
    `Price Increase Rate` = increase_rate,
    `Avg Buyer Close Rate` = avg_buyer_close_rate,
    `Avg Seller Close Rate` = avg_seller_close_rate
  ) %>%
  
  #Format the numbers
  mutate(
    Score = round(Score, 2), 
    Population = comma(Population, accuracy = 1),
    Customers = comma(Customers, accuracy = 1),
    `Median Price` = dollar(`Median Price`, accuracy = 1),
    `Price Increase Rate` = percent(`Price Increase Rate`, accuracy = 0.1),
    `Avg Buyer Close Rate` = percent(`Avg Buyer Close Rate`, accuracy = 0.1),
    `Avg Seller Close Rate` = percent(`Avg Seller Close Rate`, accuracy = 0.1)
  )

#Generate table
efficiency_table_data %>%
  kbl(caption = "Top 5 Markets by Efficiency-Weighted Score") %>%
  kable_classic_2(full_width = F, html_font = "Cambria") %>%
  row_spec(0, bold = T, background = "#f2f2f2") %>%
  # Highlight the Rank and Score columns with distinct efficiency colors
  column_spec(1, bold = T, background = "#09223d", color = "white") %>% 
  column_spec(2, bold = T, color = "#44b2ef")                      
```

After accounting for population and high property values, the top five markets 
are Grand Rapids, Raleigh, Buffalo, San Jose, and Louisville. These markets have 
the best balance between liquidity (they don't spend a lot of time on the 
market), listing price (with the exception of San Jose, they present a low 
barrier to entry for a buyer), listing price increase rates, and agent closing 
rates, and thus represent the best opportunity to maximize return on investment 
(ROI).

### 2.1.3 Market Data Trend 📈📉
Markets with the highest turnover are the least likely to face spiking listing 
prices. This suggests that the highest volume markets are highly mature and 
efficient, so buyers and sellers face less volatility in these markets.

In contrast, low turnover markets have the highest price volatility. These 
markets may be experiencing "boom" conditions due to a sudden lack of supply or
short-term enthusiasm, making them riskier. An investor is most likely to find 
consistent, steady returns - rather than speculative gains - in the high-turnover, low-volatility market.

```{r market_trend_chart, echo=FALSE, message=FALSE}
market_trend_chart <- market_ranking_data %>%
  filter(!is.na(turnover_rate) & !is.na(increase_rate) & !is.na(days_on_market_median)) %>%
  
 ggplot(aes(
    x = turnover_rate,
    y = increase_rate,
    color = days_on_market_median #Use DoM for color to see where fast/slow markets fall
  )) +
  
  #Linear trendline
  geom_smooth(
    method = "lm", 
    se = TRUE, 
    color = "black", 
    linetype = "dashed", 
    linewidth = 0.8
  ) +
  
  geom_point(aes(size = 3), alpha = 0.7) + 
  
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  
  labs(
    title = "Market Trend: High Turnover Yields Lower Price Appreciation",
    subtitle = "Inverse relationship indicates market maturity and efficiency",
    x = "Turnover Rate", 
    y = "Price Increase Rate"
  ) +
  
  #Color scheme
  scale_color_gradient(
    low = "#09223d", 
    high = "#44b2ef", 
    name = "Days on Market"
  ) + 

  guides(size = "none") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold")
  )

ggplotly(
  market_trend_chart, 
  tooltip = c("x", "y", "color")
) %>%
  style(
    text = paste0(
      "<b>Market ID:</b> ", market_ranking_data$market_id, "<br>",
      "<b>Turnover Rate:</b> ", round(market_ranking_data$turnover_rate * 100, 1), "%", "<br>",
      "<b>Price Increase Rate:</b> ", round(market_ranking_data$increase_rate * 100, 1), "%", "<br>",
      "<b>Days on Market:</b> ", round(market_ranking_data$days_on_market_median, 0)
    ),
    hoverinfo = "text",
  colorbar = list(
      orientation = "h",        
      x = 0.5,                  
      y = -0.15,                
      xanchor = "center",
      yanchor = "top",
      len = 0.7,                
      title = list(
        text = "<b>Days on Market</b>",
        side = "top",
        font = list(
          weight = "bold"
      )
    )
  )
) %>%
  layout(
    title = list(
      xanchor = "left", 
      x = 0
    ),
    margin = list(l = 50) 
  )
```
---

## 2.2 Best Agents️️ 🏋

What makes an agent the best is just as nuanced as what makes a market the best.
Let's look at some of the key factors! To determine the best agent in each
market, we'll look at their number of ratings, their average rating, how many
deals they've closed, and the percentage of leads they've converted to sales.

```{r agent_scoring, include=FALSE}
#Calculate scoring columns
agents <- agents %>%
  mutate(
    #Weighted rating
    customer_reviews_2024 = replace_na(customer_reviews_2024, 0),
    avg_rating_2024 = replace_na(avg_rating_2024, 0),
    weighted_rating = avg_rating_2024 * sqrt(customer_reviews_2024),
    
    #Total raw closes
    total_raw_closes = buyer_closes_2024 + seller_closes_2024
  )

#Select metrics and filter out inactive agents
agent_ranking_metrics <- agents %>%
  filter(total_raw_closes > 0 | customer_reviews_2024 > 0) %>%
  select(
    agent_id, market_id,
    total_raw_closes, buyer_close_rate, seller_close_rate, 
    weighted_rating, avg_rating_2024, customer_reviews_2024
  )

#Standardize and rename Z-scores 
agent_ranking_zscores <- agent_ranking_metrics %>%
  select(-market_id) %>% 
  
  #Standardize columns
  mutate(across(c(total_raw_closes:customer_reviews_2024), 
                scale, .names = "Z_{.col}")) %>%
  
  #Keep only agent_id and the Z_ columns
  select(agent_id, starts_with("Z_")) 

#Join original metrics with Z-scores using agent_id
agent_ranking_final <- agent_ranking_metrics %>%
  left_join(agent_ranking_zscores, by = "agent_id") %>%
  
  # Calculate the weighted score
  mutate(
    performance_score = 
      (Z_buyer_close_rate * 1.0) + #direct skill at converting leads to sales
      (Z_seller_close_rate * 1.0) + #direct skill at converting leads to sales
      (Z_total_raw_closes * 0.75) + #weighted down to mitigate population bias
      (Z_weighted_rating * 1.25) + #most reliable measure, accounts for small sample size bias
      (Z_avg_rating_2024 * 0.5) + #weighted down to mitigate small sample size bias
      (Z_customer_reviews_2024 * 0.5) #measures customer engagement, an indirect measure
  )

top_agent_per_market <- agent_ranking_final %>%
  group_by(market_id) %>%
  #Select the agent with the highest score in each market
  slice_max(performance_score, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  arrange(desc(performance_score)) %>%
  #Select relevant columns
  select(market_id, agent_id, performance_score, 
         total_raw_closes, buyer_close_rate, seller_close_rate, 
         weighted_rating, avg_rating_2024, customer_reviews_2024)
```

```{r agent_table}
top_10_agents_table <- top_agent_per_market %>%
 
   head(10) %>% 
  
  # Select the most important metrics
  select(
    Rank = performance_score, 
    Market = market_id,
    Agent = agent_id,
    Score = performance_score,
    `Weighted Rating` = weighted_rating,
    `Total Closes` = total_raw_closes,
    `Buyer Rate` = buyer_close_rate,
    `Seller Rate` = seller_close_rate,
    `Total Reviews` = customer_reviews_2024
  ) %>%
  
  # Format the numbers
  mutate(
    Rank = row_number(), 
    Score = round(Score, 2),
    `Weighted Rating` = round(`Weighted Rating`, 2),
    `Total Closes` = comma(`Total Closes`, accuracy = 1),
    `Buyer Rate` = percent(`Buyer Rate`, accuracy = 0.1),
    `Seller Rate` = percent(`Seller Rate`, accuracy = 0.1),
    `Total Reviews` = comma(`Total Reviews`, accuracy = 1)
  ) %>%
  
  # Reorder columns to place rank first
  select(Rank, Market, Agent, everything())

# Generate the HTML table
top_10_agents_table %>%
  kbl(caption = "Top 10 Agents by Composite Performance Score (Best Agent from Each Market)") %>%
  kable_classic_2(full_width = F, html_font = "Cambria") %>%
  # Highlight the Rank and Score
  column_spec(1, bold = T, background = "#09223d", color = "white") %>%
  column_spec(4, bold = T, color = "#44b2ef") %>% # Score
  column_spec(5, bold = T, color = "#44b2ef") %>% # Weighted Rating
  row_spec(0, bold = T, background = "#f2f2f2")
```
<small>The table above highlights the top 10 agents overall. For a complete 
listing of the best agent in every market analyzed, the complete, ranked list 
can be downloaded [here](best_agents_per_market.csv).</small>
 
```{r full_agent_ranking, include=FALSE}
write.csv(top_agent_per_market, 
          file = "best_agents_per_market.csv", 
          row.names = FALSE)
```
 
So what makes these agents the best? 

Though the number of customer reviews was weighted down in this analysis, it's 
clear that customer engagement matters! All but one of the top agents had higher 
than the average number of customer reviews. The average number of customer
reviews per agent overall is 4.6, while the top agents had an average of 8.9 
reviews - almost double the customer engagement rate! 

```{r agent_success_customer_engagement, include=FALSE}
mean_customer_reviews_2024 <- mean(agents$customer_reviews_2024)
print(mean_customer_reviews_2024)
mean_top_agent_customer_reviews_2024 <- 
  mean(top_agent_per_market$customer_reviews_2024)
print(mean_top_agent_customer_reviews_2024)

percent_customer_review_2024 <- (sum(top_agent_per_market$customer_reviews_2024 > mean_customer_reviews_2024)/ nrow(top_agent_per_market)) * 100
print(percent_customer_review_2024)
```

<br>

```{r agent_success_customer_engagement_chart, echo=FALSE}
colors <- c('#44b2ef', '#09223d')

customer_review_average_compare <- plot_ly(
  x = c("<b>Average Agents</b>", 
        "<b>Top Agents</b>"),
  y = c(mean_customer_reviews_2024, mean_top_agent_customer_reviews_2024),
  name = "<b>Average Number of Customer Reviews</b>",
  type = "bar",
  text = sprintf("<b>%.1f reviews</b>", c(mean_customer_reviews_2024, mean_top_agent_customer_reviews_2024)), 
  textposition = 'auto',
  hovertemplate = "%{x} have %{y:.1f} customer reviews<extra></extra>",
         marker = list(color = colors,line = list(color = 'rgb(8,48,107)', 
                                                  width = 1.5))
)
customer_review_average_compare <- customer_review_average_compare %>% layout(title = "<b>Average Number of Customer Reviews Compared to Top Agents</b>",
         xaxis = list(title = ""),
         yaxis = list(title = "")
)

customer_review_average_compare

customer_review_average_percent <- plot_ly(labels = c("Top Agents Who Exceed Average Number of Ratings", "Other"), values = c(percent_customer_review_2024, (100-percent_customer_review_2024)), 
                                           type = 'pie',
                                           marker = list(
        colors = c('#44b2ef', '#09223d')
    ),
    textinfo = 'label+percent',
                                         texttemplate = '<b>%{percent}</b>',
    hovertemplate = "<b>%{label}:</b> %{percent}<extra></extra>")
customer_review_average_percent <- customer_review_average_percent %>% 
layout(title = '<b>Top Agents with Higher than Average Number of Reviews</b>',
       margin = list(
               t = 125 
           ),
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
       legend = list(
    orientation = "h", 
    x = 0.5,             
    y = -0.2,
    xanchor = "center"))

customer_review_average_percent
```
<br>

The best agents are also able to maintain a steady volume of customers. Volume
isn't the main driver, but more (engaged!) customers means more sales! All but 
two of the top agents closed as many or more sales than average. On average, top
agents closed 50% more sales than the average agent!

```{r agent_success_volume, include=FALSE}
mean_total_raw_closes <- mean(agents$total_raw_closes)
print(mean_total_raw_closes)

mean_top_agent_total_raw_closes <- mean(top_agent_per_market$total_raw_closes)
print(mean_top_agent_total_raw_closes)

percent_total_raw_closes <- (sum(top_agent_per_market$total_raw_closes > mean_total_raw_closes)/ nrow(top_agent_per_market)) * 100
print(percent_total_raw_closes)
```

<br>

```{r agent_success_volume_chart}
colors <- c('#44b2ef', '#09223d')

total_raw_closes_compare <- plot_ly(
  x = c("<b>Average Agents</b>", 
        "<b>Top Agents</b>"),
  y = c(mean_total_raw_closes, mean_top_agent_total_raw_closes),
  name = "Average Number of Closed Deals",
  type = "bar",
  text = sprintf("<b>%.1f closed deals</b>", c(mean_total_raw_closes, mean_top_agent_total_raw_closes)), 
  textposition = 'auto',
  hovertemplate = "%{x} closed %{y:.1f} deals<extra></extra>",
         marker = list(color = colors,line = list(color = 'rgb(8,48,107)', 
                                                  width = 1.5))
)
total_raw_closes_compare <- total_raw_closes_compare %>% layout(title = 
                                                                  "<b>Average Number of Closed Deals Compared to Top Agents</b>",
         xaxis = list(title = ""),
         yaxis = list(title = "")
)

total_raw_closes_compare

total_raw_closes_percent <- plot_ly(labels = c("Top Agents Who Exceed Average Number of Closed Deals", "Other"), values = c(percent_total_raw_closes, (100-percent_total_raw_closes)), 
                                           type = 'pie',
                                           marker = list(
        colors = c('#44b2ef', '#09223d')
    ),
    textinfo = 'label+percent',
                                         texttemplate = '<b>%{percent}</b>',
    hovertemplate = "<b>%{label}:</b> %{percent}<extra></extra>")
total_raw_closes_percent <- total_raw_closes_percent %>% 
layout(
  title = '<b>Top Agents with Higher than Average Number of Closed Deals</b>',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
  margin = list(
  t=125),
       legend = list(
         orientation = "h", 
    x = 0.5,             
    y = -0.2,
    xanchor = "center"))

total_raw_closes_percent
```

Over half of top agents convert more leads into sales than average. The best agents
are able to close 25-35% more deals than average!

<br>

```{r agent_success_close, include=FALSE}
mean_buyer_close_rate <- mean(agents$buyer_close_rate)
mean_seller_close_rate <- mean(agents$seller_close_rate)

print(mean_buyer_close_rate)
print(mean_seller_close_rate)

mean_top_agent_buyer_close_rate <- mean(top_agent_per_market$buyer_close_rate)
mean_top_agent_seller_close_rate <- mean(top_agent_per_market$seller_close_rate)

print(mean_top_agent_buyer_close_rate)
print(mean_top_agent_seller_close_rate)

percent_buyer_close_rate <- (sum(top_agent_per_market$buyer_close_rate > mean_buyer_close_rate)/ nrow(top_agent_per_market)) * 100
print(percent_buyer_close_rate)
percent_seller_close_rate <- (sum(top_agent_per_market$seller_close_rate > mean_seller_close_rate)/nrow(top_agent_per_market)) * 100
print(percent_seller_close_rate)
```

```{r agent_success_close_chart}
Categories <- c("Average Agents", "Top Agents")
Buyer_Rate = c(mean_buyer_close_rate, mean_top_agent_buyer_close_rate)
Seller_Rate = c(mean_seller_close_rate, mean_top_agent_seller_close_rate)
data <- data.frame(Categories, Buyer_Rate, Seller_Rate)

# Define colors
colors <- c('#44b2ef', '#09223d')

#Define text labels for the bars
buyer_text_labels <- paste0("<b>",round(data$Buyer_Rate * 100, 1), "%</b>")
seller_text_labels <- paste0("<b>", round(data$Seller_Rate * 100, 1), "%</b>")

total_close_rate_compare <- plot_ly(data,
    x = ~Categories,
    y = ~Buyer_Rate,
    name = "Buyer Close Rate",
    type = "bar",
    hovertemplate = "<b>%{x}<br>Buyer Rate:</b> %{y:.1%}<extra></extra>",
    text = buyer_text_labels,
    textposition = 'outside',
    textfont = list(weight = 'bold', size = 12),
    marker = list(
        color = colors[1],
        line = list(color = 'rgb(8,48,107)', width = 1.5)
    )
)
total_close_rate_compare <- total_close_rate_compare %>%
    add_trace(
        y = ~Seller_Rate,
        name = 'Seller Close Rate',
        hovertemplate = "<b>%{x}<br>Seller Rate:</b> %{y:.1%}<extra></extra>",
        text = seller_text_labels,
        textposition = 'outside',
        textfont = list(weight = 'bold', size = 12),
        marker = list(
            color = colors[2],
            line = list(color = 'rgb(8,48,107)', width = 1.5)
        )
    )

total_close_rate_compare <- total_close_rate_compare %>%
    layout(
        title = "<b>Average Close Rate Compared to Top Agents</b>",
        xaxis = list(title = "", showticklabels = FALSE),
        yaxis = list(
            title = '<b>Close Rate</b>',
            tickformat = '.0%'),
        barmode = 'group',
legend = list(
            orientation = 'h',
            y = -0.2, 
            x = 0.5, 
            xanchor = 'center',
            font = list(weight = 'bold')
        ),
annotations = list(
            list(
                x = "Average Agents", 
                y = 0, # Position below the bar, adjust if needed
                text = "<b>Average Agents</b>", 
                showarrow = FALSE, 
                xref = "x", 
                yref = "paper", 
                yanchor = "top",
                yshift = -30 # Adjust vertical position 
            ),
            list(
                x = "Top Agents", 
                y = 0, 
                text = "<b>Top Agents</b>", 
                showarrow = FALSE, 
                xref = "x", 
                yref = "paper",
                yanchor = "top",
                yshift = -30 # Adjust vertical position
            )
        )
    )

total_close_rate_compare

colors <- c('#229fe6', '#2a4a6a')

percent_buyer_close_rate_chart <- plot_ly(
    labels = c("Top Agents Exceeding Average Buyer Close Rate", "Other"),
    values = c(percent_buyer_close_rate, (100 - percent_buyer_close_rate)),
    type = 'pie',
    marker = list(colors = colors),
    hovertemplate = "<b>%{label}:</b> %{percent}<extra></extra>",
    domain = list(x = c(0, 0.45)),
    name = "Buyer Rate",
    textinfo = 'percent',
    textfont = list(weight = 'bold', size = 12),
    texttemplate = '<b>%{percent}</b>',
    textposition = 'inside'
) %>% layout(title = '<b>Buyer Close Rate</b>')

percent_seller_close_rate_chart <- plot_ly(
    labels = c("Top Agents Exceeding Average Seller CloseRate", "Other"),
    values = c(percent_seller_close_rate, (100 - percent_seller_close_rate)),
    type = 'pie',
    marker = list(colors = colors),
    hovertemplate = "<b>%{label}:</b> %{percent}<extra></extra>",
    domain = list(x = c(0.55, 1)),
    name = "Seller Rate",
    textinfo = 'percent',
    textfont = list(weight = 'bold', size = 12),
    texttemplate = '<b>%{percent}</b>',
    textposition = 'inside'
) %>% layout(title = '<b>Seller Close Rate</b>')

combined_pie_charts <- subplot(
    percent_buyer_close_rate_chart,
    percent_seller_close_rate_chart,
    nrows = 1,
    margin = 0.05
)

combined_pie_charts <- combined_pie_charts %>%
    layout(
        title = '<b>Percentage of Top Agents Exceeding Average Close Rates</b>',
        showlegend = FALSE
    )

combined_pie_charts
```

The bottom line? The best agents are the ones who can effectively engage with
a high volume of customers without sacrificing quality. More customers means
more opportunities, but only if you can keep them happy!

## 2.3 Area for Growth 🪴
It's tempting to want to stick with what you know - lots of people and an
expensive market means lots of opportunity for high-commission sales! However,
there's opportunity beyond wealthy, densely populated markets like New York City 
and Los Angeles.

As discussed above, New York City and Los Angeles are two of the most successful
markets right now. However, *highest value* doesn't always mean *most efficient*.
Investing more resources in high-efficiency markets, like Grand Rapids, MI or
Raleigh, NC, could lead to higher turnover, the opportunity for more - if 
smaller - sales, and the chance to connect with more people by being able to
close more deals in a shorter amount of time.

As a case study, let's compare New York City (a high value market) to Grand 
Rapids (a highly efficient market).

```{r case_study}
grand_rapids_price <- markets %>%
  filter(market_id == "grand_rapids_mi") %>%
  pull(price_closed_median)

grand_rapids_dom <- markets %>%
  filter(market_id == "grand_rapids_mi") %>%
  pull(days_on_market_median)

grand_rapids_turnover <- markets %>%
  filter(market_id == "grand_rapids_mi") %>%
  pull(turnover_rate)

new_york_price <- markets %>%
  filter(market_id == "new_york_ny") %>%
  pull(price_closed_median)

new_york_dom <- markets %>%
  filter(market_id == "new_york_ny") %>%
  pull(days_on_market_median)

new_york_turnover <- markets %>%
  filter(market_id == "new_york_ny") %>%
  pull(turnover_rate)

data_case_study <- tibble(
  Market = c("<b>New York, NY</b>", "<b>Grand Rapids, MI</b>"),
  Price_Closed_Median = c(new_york_price, grand_rapids_price),
  Days_On_Market_Median = c(new_york_dom, grand_rapids_dom),
  Turnover_Rate = c(new_york_turnover, grand_rapids_turnover)
)

#Raw Market Performance Comparison
data_long_case_study <- data_case_study %>%
  pivot_longer(
    cols = c(Price_Closed_Median, Days_On_Market_Median, Turnover_Rate),
    names_to = "Metric",
    values_to = "Value"
  ) %>%
  mutate(
    Metric_Type = case_when(
      Metric == "Price_Closed_Median" ~ "Price Closed Median ($)",
      Metric == "Days_On_Market_Median" ~ "Days On Market Median (Days)",
      Metric == "Turnover_Rate" ~ "Turnover Rate",
      TRUE ~ Metric
    )
  )

#Define conditional formatting for the hover text
data_long_case_study_formatted <- data_long_case_study %>%
  mutate(
    Formatted_Value_Hover = case_when(
      Metric == "Price_Closed_Median" ~ paste0("$", format(round(Value, 0), big.mark = ",", scientific = FALSE, trim = TRUE)),
      Metric == "Days_On_Market_Median" ~ paste0(format(round(Value, 0), big.mark = ",", scientific = FALSE, trim = TRUE), " days"),
      Metric == "Turnover_Rate" ~ paste0(format(round(Value * 100, 2), big.mark = ",", scientific = FALSE, trim = TRUE, nsmall = 2), "%"),
      TRUE ~ as.character(Value)
    )
  )

raw_market_comparison <- ggplot(data_long_case_study_formatted, aes(x = Market, y = Value, fill = Market, text = paste0("<b>", Market, "</b>\n", 
                         "<b>", Metric_Type, ": </b>", 
                         Formatted_Value_Hover))) +
  geom_col(position = "dodge") +
  geom_text(aes(y = Value * 0.09,
                label = paste0("<b>", case_when(
      Metric == "Price_Closed_Median" ~ format(round(Value, 0), big.mark = ",", scientific = FALSE, trim = TRUE),
      Metric == "Days_On_Market_Median" ~ format(round(Value, 0), big.mark = ",", scientific = FALSE, trim = TRUE),
      Metric == "Turnover_Rate" ~ paste0(format(round(Value * 100, 2), big.mark = ",", scientific = FALSE, trim = TRUE, nsmall = 2), "%"),
      TRUE ~ format(Value)
    ), "</b>")),
    position = position_dodge(width = 0.9),
    vjust = 1, 
    size = 3,
    fontface = "bold"
  ) +
  scale_y_continuous(labels = scales::comma) +
  facet_wrap(~ Metric_Type, scales = "free_y", ncol = 2) +
  labs(
    title = "<b>Raw Market Performance Comparison (NY vs GR)</b>",
    y = "",
    x = ""
  ) +
  scale_fill_manual(values = c("<b>New York, NY</b>" = "#2a4a6a", "<b>Grand Rapids, MI</b>" = "#229fe6")) +
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 0, hjust = 0.5, face = "bold"),
        axis.text.y = element_text(face = "bold"),
        plot.title = element_text(face = "bold"),
    strip.text = element_text(face = "bold"),
    axis.title = element_text(face = "bold"))

chart1 <- ggplotly(raw_market_comparison, tooltip = "text") %>%
  layout(title = list(text = "<b>Raw Market Performance Comparison (NY vs GR)</b>"))

chart1
```

New York requires 121% more capital for a transaction than Grand Rapids, 
significantly hindering liquidity and sales velocity. A property in New York
will also sit four times longer on average than a property in Grand Rapids. This
means that a salesperson could theoretically sell four properties in Grand 
Rapids for the same amount of time and energy it takes to sell just one property
in New York - based on median sales prices at close, that's a gross transaction 
value difference of $637,000 in Grand Rapids over New York! With a lower barrier 
to entry and higher turnover, Grand Rapids has 88% higher market activity than 
New York. Investing more resources into smaller, highly efficient markets like 
Grand Rapids will lead to a higher return on investment (ROI) than focusing 
exclusively on large markets like New York.

---
# 3. Summary 📋

## 3.1 What If There Was More Time? ⌚
The analysis could be improved by moving beyond static rate comparisons and into 
predictive modeling and geospatial segmentation. There are three primary ways I
would improve the above analysis, given the time and resources.

### 3.1.1 Predictive Agent Performance Modeling
Instead of only comparing current averages, I would build a model to predict the 
future success (e.g. close rate, tenure, revenue) of a new agent entering a 
specific market. This would separate agent quality from market quality. To build 
a successful model, I would need agent profile data to learn historical agent 
demographics, training scores, prior industry experience, and agent-specific 
lead sources. I would also use lead quality metrics to tag leads by source 
(e.g., paid digital, referral, organic search) and initial qualification score. 
Lastly, I would use a model output to create a propensity score for each market 
and agent profile, showing the predicted probability of an agent achieving a 
top-quartile close rate within 12 months.

### 3.1.2. Geospatial Analysis (Micro-Market Segmentation)
The current analysis treats each city as a monolithic market. In reality, a 
massive city like New York is composed of highly distinct micro-markets (e.g. 
Manhattan vs. Brooklyn vs. Long Island). With more time I could use geospatial
boundary data to create detailed maps of county, zip code, or neighborhood
service areas for each transaction. I could then also add more specific local 
economic indicators like median income, population density, average home sale
price, and inventory days on market to those micro-markets. With these data, I
could dive much deeper than the current analysis and visualize data like the top
ten highest-ROI zip codes, which would enable us to find high efficiency zip
codes and focus sales to maximize local ROI.

### 3.1.3. Time-Series Analysis
The current metrics are static averages taken from a snapshot of data. To 
understand how agent performance, customer saturation, and growth rates 
fluctuate over time, I could use historical data to track transaction
volumes, lead flow, agent performance, etc and analyze how those metrics have
changed over the years. This would provides the necessary context to distinguish 
a temporary spike from a long-term trend. Having time-stamped external factors 
like local interest rate data, state or local regulatory changes, or major 
economic events would also help make the analysis more robust; for example, 
these data could help explain the outliers in listing price decreases in 
Columbus and Indianapolis seen above. A time-series analysis could point to any 
trends in seasonality or other peak performance windows.

## 3.2 How did this report come together? 📜
The overall process followed the standard data analysis workflow: data cleaning
and preparation, analytical modeling, and communication. The entire project was 
executed primarily using R, leveraging packages like tidyverse for data 
manipulation and ggplot2, plotly, and kableExtra for visualization and 
reporting.

If the data isn't clean, all of my analyses could give me wildly different 
results than expected. When I clean data, I make sure to address a few key 
areas: data type consistency, missingness, potential spelling errors or typos 
that could lead to aggregation errors, logical tests (e.g. dates shouldn't 
be in the future), and outliers. One challenge in data cleaning is deciding how 
to handle data that doesn't pass my checks. Sometimes data can be imputed (e.g. 
a rating score can be imputed with the median for a dataset like this one with 
low missingness because it won't introduce skew). Sometimes missing values can 
be safely deleted. Sometimes outliers are an obvious data entry error, but 
without access to the source data, that can be hard to verify. Ultimately, I 
have to use my critical thinking skills to make the best decision for each 
dataset to maintain data integrity.

After the data is clean, it's on to the fun part! To perform the analyses, I
think critically about what is being asked and the best metric to answer those
questions. For this project I primarily used R, but I have also used Python, 
Excel, SQL, Tableau, and Infogram in other projects.

I started by creating derived metrics like avg_buyer_close_rate and
customer_saturation_rate by grouping the raw agents and customers data by 
market_id. This established the foundation for subsequent comparisons.

To compare markets and agents fairly, I used the scale() function in R to 
convert raw metrics (like median price, close rates, etc.) into Z-scores. This 
process, called normalization, controlled for the sheer size and wealth of major 
cities (like New York or Los Angeles), ensuring the resulting scores were based 
on performance relative to the average, not just the absolute numbers.

I then created the final Composite Performance Scores for both markets and 
agents by assigning weights to these Z-scores. For example, in the Efficiency 
Market Analysis, I intentionally gave lower weight to the Z-score for 
price_closed_median and higher weight to close rates to favor operational 
efficiency over market wealth. For the Agent Analysis, I gave the highest weight 
to the weighted_rating to favor agents with proven customer satisfaction.

Finally, I used the ggplot2, plotly, and kableExtra packages to create the 
visualizations and clean tables, ensuring that the highest performers were 
easily identifiable.

One of the most significant challenges was the inherent multi-collinearity and 
size bias in the raw data. My first attempt at ranking the markets showed the 
largest cities (like New York and Los Angeles) at the top. This analysis was ok, 
but it seemed shallow and pointed to the fact that data based on population size 
had not been accounted for well enough, despite previous attempts at 
normalization. This confirmed that raw size was overpowering other metrics.

I addressed the challenge by enforcing the standardization step and then 
adjusting the weights (e.g., reducing the weight of volume metrics and 
increasing the weight of efficiency metrics) to successfully decouple 
operational success from raw market size.

Analyses don't mean much if no one can understand them. After the analyses, I
clean up the report. I make sure the report itself is formatted and easy to
read. I put the finishing touches on all the visualizations to make sure the
data I want to stand out really pops. Graphic design and intentional formatting
can go a long way to enhancing communication, especially for subjects that can
otherwise be dry or tricky to understand.

Throughout the whole process, I make sure my work is saved and backed up. I use
Github to version control my work, as well as to provide a secondary backup.
